---
title: "Class6"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(DBI)
library(RSQLite)
```

Sort weekdate
order_day <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
mutate(Weekday = weekdays(Date)) %>% mutate(Weekday = factor(Weekday, order_day))
```{r}
weekdays(Sys.Date())
```

# SL SQL
Connect to the database, list the tables and figure out how they relate to eachother.

```{r}
con <- dbConnect(SQLite(), "Class_files/sthlm_metro.sqlite")
tables <- dbListTables(con)
#lapply(tables, dbReadTable, conn = con)
```

Query for the LineName from table Line where LineNumber is 18.

```{r}
dbGetQuery(con, "SELECT StationName, count(PlatformNumber) AS Number 
           FROM Platform GROUP BY StationName ORDER BY Number DESC")
```

Query for the five most southern StationName, where position is measured as the average Latitude of its PlatformNumber.

```{r}
dbGetQuery(con, "SELECT StationName, avg(Latitude) AS avg_Latitude FROM Platform 
           GROUP BY StationName ORDER BY avg_Latitude DESC LIMIT 5")
```

Query for the number of stations on LineNumber 18.

```{r}
dbGetQuery(con, "SELECT LineNumber, count(DISTINCT StationName) AS Number
           FROM LinePlatform LEFT JOIN Platform 
           ON LinePlatform.PlatformNumber = Platform.PlatformNumber
           WHERE LineNumber = 18")
```

Query for all StationName on LineNumber 18 in alphabetical order.

```{r}
dbGetQuery(con, "SELECT DISTINCT StationName FROM LinePlatform LEFT JOIN Platform 
           ON LinePlatform.PlatformNumber = Platform.PlatformNumber
           WHERE LineNumber = 18")
dbDisconnect(con)
```


# Pokemon SQL

```{r}
con <- dbConnect(SQLite(), "Class_files/pokemon.sqlite")
tables <- dbListTables(con)
lapply(tables, dbReadTable, conn = con)
```

Query for the average height and average weight of pokemon (pokemon table).

```{r}
dbGetQuery(con, "SELECT avg(height) AS avg_height, avg(weight) AS avg_weight
           FROM pokemon")
```

Query for the number of pokemon of less than average weight.

```{r}
dbGetQuery(con, "SELECT COUNT(*) AS number
            FROM pokemon
            WHERE weight < (SELECT avg(weight) FROM pokemon)")
```

Query for the average weight of ghost-type pokemon.

```{r}
dbGetQuery(con, "SELECT avg(pokemon.weight) AS avg_weight
           FROM pokemon 
           LEFT JOIN pokemon_types ON pokemon.id = pokemon_types.pokemon_id
           LEFT JOIN types ON pokemon_types.type_id = types.id
           WHERE types.identifier = 'ghost'")
```



```{r}
dbDisconnect(con)
```




