---
title: "Class12"
output: github_document
---
Thi Thuy Nga Nguyen

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(purrr)
library(httr)
library(rvest)
library(stringi)
library(stringr)
```

# Solve exercises in R4DS chapter 21.5.3
Write code that uses one of the map functions to: Compute the mean of every column in mtcars.

```{r}
map_dbl(mtcars, mean, na.rm = TRUE)
```

Determine the type of each column in nycflights13::flights.

```{r}
map_chr(nycflights13::flights, typeof)
```

Compute the number of unique values in each column of iris.

```{r}
iris %>%
  map(unique) %>%
  map_dbl(length)
```

Generate 10 random normals for each of $\mu$ = -10, 0, 10, 100

```{r}
map(c(-10, 0, 10, 100), rnorm, n = 10)
```

How can you create a single vector that for each column in a data frame indicates whether or not it’s a factor?

```{r}
map_lgl(iris, is.factor)

map(1:5, runif)  # 1:5 -> n

map(-2:2, rnorm, n = 5)  # -2:2 -> mean
```

Eliminate the anonymous function

```{r}
mtcars %>% 
  split(.$cyl) %>% 
  map(function(df) lm(mpg ~ wt, data = df))
mtcars %>% 
  split(.$cyl) %>%
  map(~lm(mpg ~ wt, data = .))
```

# A likelihood function

```{r}
L_exp <- function(x, lambda) prod(dexp(x, lambda))
set.seed(1)
x <- rexp(100, 1)
L_exp(x, 1)
lambda <- seq(0.5, 2, length.out = 10)
L_exp(x, lambda)
```

Use map_dbl to compute the vector $(L(\lambda_1),…,L(\lambda_{10}))$

```{r}
map_dbl(lambda, L_exp, x = x)
```


# More dramatic
Modify the function repertoir from day 12 (Class11) such that it generates a suitable result if the API call fails (e.g. due to an empty repertoir). Use it in combination with e.g. map_df in order to create a table of all sets from 1908-2017. Who is the most active director?

```{r}
repertoire <- function(year){
  html <- read_html(paste0("https://www.dramaten.se/medverkande/rollboken/?category=date&query=", year))
  get_column <- function(css) {
    html %>%
      html_nodes(css) %>%
      html_text()
  }
  df <- data.frame(Play = get_column(".play-name")[-1], 
                   Opening_night = get_column(".first-appearance")[-1],
                   Director = get_column(".director")[-1], Stage = get_column(".stage")[-1])
  df$Director <- df$Director %>%
    str_extract_all("[A-ZÄÅÖ][a-ö-A-Ö]+ [A-ZÄÅÖ][a-ö-A-Ö]+") %>%
    lapply(function(x) if(identical(x,character(0))) "" else x) %>%
    unlist()
  df <- df %>%
    mutate(Play = as.character(Play), Opening_night = as.POSIXct(Opening_night), Stage = as.character(Stage))
  return(df)
}

safe_repertoire <- safely(repertoire)
years <- 1908:2017
data <- map(years, safe_repertoire)
dfs <- transpose(data)[["result"]]
errs <- transpose(data)[["error"]]
is_ok <- map_lgl(errs, is.null)
years[!is_ok]
dramatic <- map_dfr(dfs[is_ok], as.data.frame)

dramatic %>%
  group_by(Director) %>%
  summarize(number = n()) %>%
  arrange(desc(number)) %>%
  filter(Director != "") %>%
  head(n = 1)

dramatic %>%
  mutate(Year = format(Opening_night, "%Y")) %>%
  filter(Year %in% years[!is_ok])
```


# More SHL-players

```{r}
get_player <- function(url){
  if (GET(paste0(url, "/statistics"))$status_code != 200) {
    stop("error in GET url")
  } else {
    response <- GET(paste0(url, "/statistics"))
  }
  player_xml <- content(response, "text") %>% 
    read_html(player_page)
  css <- list(".rmss_c-squad__player-header-name-h", # Namn
              ".rmss_c-squad__player-header-name-info-position", #Position
              ".rmss_c-squad__player-header-info-items-item-value") # Info
  map(css, html_nodes, x = player_xml) %>% 
    map(html_text) %>% 
    unlist() %>% 
    .[1:8] %>%
    set_names(c("Namn", "Position", "Födelsedatum", "Ålder", "Nationalitet", "Längd", "Vikt", "Skjuter")) %>%
    data.frame() %>%
    t() %>%
    data.frame() %>%
    mutate(Namn = as.character(Namn), Position = as.character(Position), 
           Födelsedatum = as.POSIXct(Födelsedatum), Ålder = as.numeric(Ålder),
           Nationalitet = as.character(Nationalitet), Skjuter = as.character(Skjuter))
}
get_player("http://www.shl.se/lag/087a-087aTQv9u__frolunda-hc/qQ9-a5b4QRqdS__ryan-lasch")
```

Construct a new function get_team, that given an url of the form "http://www.shl.se/lag/2459-2459QTs1f__djurgarden-hockey/roster" generates a  data.frame of players and informations from get_player.

```{r}
get_team <- function(url) {
  response <- GET(url)
  if (response$status_code != 200) {
    stop("link is not valid!!")
  }
  player_url <- content(response, "text") %>% 
    read_html() %>%
    html_nodes(".rmss_c-squad__team-cont-roster-group-item a") %>%
    html_attr("href") %>%
    map_chr(function(x) paste0("http://www.shl.se", x))
  map_dfr(player_url, get_player)
}
get_team("http://www.shl.se/lag/2459-2459QTs1f__djurgarden-hockey/roster") %>% 
  head()
```



