---
title: "Class 4: Exploratory data analysis"
output: github_document
---
Nga Nguyen

```{r setup}
library(ggplot2)
library(tidyverse)
library(knitr)
library(readr)
```

## Systembolaget’s assortment
Use filter to extract the groups of products c("Vitt vin", "Rött vin", "Rosévin", "Mousserande vin") of vintage 2011-2018. 

```{r, Systembolaget’s assortment}
Sortiment_hela <- read_csv("Class_files/systembolaget2018-10-08.csv")
#str(Sortiment_hela)

mydata <- Sortiment_hela %>%
  filter(Varugrupp %in% c("Vitt vin", "Rött vin", "Rosévin", "Mousserande vin"), Argang %in% c(2011:2018))
```

ggplot with aes(x = Argang), geom_bar()

```{r}
ggplot(mydata, aes(x = Argang)) + geom_bar()
```

ggplot with aes(x = Argang), geom_bar() and facet_wrap(~ Varugrupp) (try adding scale = "free_y" to facet_wrap)

```{r}
ggplot(mydata, aes(x = Argang)) + geom_bar() + facet_wrap(~ Varugrupp, scale = "free_y")
```

ggplot with aes(x = Argang, fill = Varugrupp) and
geom_bar()

```{r}
ggplot(mydata, aes(x = Argang, fill = Varugrupp)) + geom_bar()
```

geom_bar(position = "dodge")

```{r}
ggplot(mydata, aes(x = Argang, fill = Varugrupp)) + geom_bar(position = "dodge")
```

geom_bar(position = "fill")

```{r}
ggplot(mydata, aes(x = Argang, fill = Varugrupp)) + geom_bar(position = "fill")
```

Recreate the following plot (Red wines in the regular range) 

```{r}
mydata %>%
  filter(Varugrupp == "Rött vin") %>%
  ggplot(aes(x = as.factor(Argang), y = PrisPerLiter)) + geom_boxplot() + ggtitle("Red wines")
```

Make a box_plot of PrisPerLiter on the log-scale,with x = Varugrupp. Try coord_flip to improve readability.

```{r}
ggplot(mydata, aes(x = as.factor(Varugrupp), y = PrisPerLiter)) + geom_boxplot() + 
  scale_y_log10() + coord_flip()
```

## Winter medals
The following code transforms the medals data to “long” format (more about this next time!) which is easier to work with in ggplot:

```{r, Winter medals}
medal_long <- read_csv("Class_files/Winter_medals2018-09-26.csv") %>% 
  select(-Total) %>% 
  gather(Denomination, Number, c("Gold", "Silver", "Bronze"))
```

Check the result with glimpse(medal_long). Use group_by and summarise in order to aggregate the total number of medals of each denomination (Gold/Silver/Bronze) for each country. Illustrate the relative proportions of denominations, e.g. by geom_bar with  stat = "identity and position = "fill".

```{r}
glimpse(medal_long)

medal_long %>%
  group_by(Denomination, Country) %>%
  summarize(Number.Denomination = sum(Number)) %>%
  ggplot(aes(x = as.factor(Country), y = Number.Denomination, fill = Denomination)) + 
  geom_bar(stat = "identity", position = "fill") +
  theme(axis.text.x = element_text(angle = 90))
```

## First math course
The file Class_files/MM2001_results.csv contains the age, sex, and grade on course Matematik I (MM2001) of 3201 students aged 18-40 years. An NA in the grade column means that the student has been registered but not yet completed the course.

Use ggplot to explore relations between the variables.

```{r, First math course}
Course_result <- read_csv("Class_files/MM2001_results.csv")

ggplot(Course_result, aes(x = Age, fill = Sex)) + geom_bar(position = "fill")
ggplot(Course_result, aes(x = Age, fill = Grade)) + geom_bar(position = "fill")
ggplot(Course_result, aes(x = Sex, fill = Grade)) + geom_bar(position = "fill")
ggplot(Course_result, aes(x = Grade, fill = Sex)) + geom_bar(position = "fill")
```

